---
title: "Site-specific Mortality Report"
author: "Australian and New Zealand Hip Fracture Registry"
date: last-modified
date-format: "DD/MM/YYYY"
format: 
    html:
        page-layout: full
        toc: false
        embed-resources: true
execute: 
  echo: false
  warning: false
  message: false
params:
  site_id: "RAH"

---


```{r}
#| results: asis

# kableExtra's bundled kePrint.js relies on jQuery-based Bootstrap 3/4
# popover initialization (`$(...).popover()`), which doesn't exist under
# Quarto's Bootstrap 5 JS. Initialize popovers with the native Bootstrap 5
# API instead.
htmltools::tags$script(htmltools::HTML(
  "document.addEventListener('DOMContentLoaded', function () {
     document.querySelectorAll('[data-toggle=\"popover\"]').forEach(function (el) {
       new bootstrap.Popover(el, {
         trigger: 'hover',
         html: true,
         container: 'body',
         placement: el.getAttribute('data-placement') || 'right',
         content: el.getAttribute('data-content')
       });
     });
   });"
))
```


```{r pkg}
library(targets)
library(tarchetypes)
library(tidyverse)
library(gt)
library(gtsummary)
library(kableExtra)
library(plotly)
```


```{r data}
tar_load(raw_data, store = here::here("_targets"))
tar_load(analysis_data, store = here::here("_targets"))
tar_load(hoscode_data, store = here::here("_targets"))

analysis_data <- labelled::set_variable_labels(
  analysis_data,
  mort30d = "30-day mortality",
  mort365d = "365-day mortality"
)
```



# `r hoscode_data$h_name[hoscode_data$report_id == params$site_id]` 

## Number of records contributed to mortality reporting


```{r tab_data_contribution}

n_raw <- raw_data |>
  filter(
    ahoscode == hoscode_data$ahoscode[hoscode_data$report_id == params$site_id]
  ) |>
  mutate(report_year = year(start_date)) |>
  group_by(report_year) |>
  tally(name = "n_raw")

n_analysis <- analysis_data |>
  filter(report_id == params$site_id) |>
  group_by(report_year) |>
  tally(name = "n_analysis")

tbl_contribution <- full_join(n_analysis, n_raw, by = "report_year") |>
  arrange(report_year) |>
  mutate(
    text = paste0(
      n_analysis,
      " / ",
      n_raw,
      " (",
      scales::percent(n_analysis / n_raw, accuracy = 1),
      ")"
    )
  ) |>
  select(report_year, text) |>
  pivot_wider(names_from = report_year, values_from = text)

tbl_contribution |>
  kbl(format = "html", escape = FALSE)
```

*Data presented as the number of records included in the analysis / the number of records entered into the registry (%)*

## Patient Demographic Characteristics

::: {.callout-tip}
## Tip
Hover over the cells in the table below to see the national reference values.
:::


```{r site_data}
site_data <- analysis_data |>
  filter(report_id == params$site_id) |>
  select(
    report_year,
    age,
    sex_2l,
    asa_nhfd,
    walk_nhfd,
    ftype_nhfd,
    uresidence_nhfd,
    mort30d,
    mort365d
  ) |>
  mutate(across(
    c(sex_2l, asa_nhfd, walk_nhfd, ftype_nhfd, uresidence_nhfd),
    ~ forcats::fct_na_value_to_level(.x, level = "Missing")
  ))
```

```{r ref_data}
ref_data <-
  analysis_data |>
  filter(report_year %in% unique(site_data$report_year)) |>
  select(
    report_year,
    age,
    sex_2l,
    asa_nhfd,
    walk_nhfd,
    ftype_nhfd,
    uresidence_nhfd,
    mort30d,
    mort365d
  ) |>
  mutate(across(
    c(sex_2l, asa_nhfd, walk_nhfd, ftype_nhfd, uresidence_nhfd),
    ~ forcats::fct_na_value_to_level(.x, level = "Missing")
  )) |>
  tbl_summary(
    by = report_year,
    missing = "no",
    digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
    statistic = list(
      all_categorical() ~ "{n} ({p}%)",
      all_continuous() ~ "{mean} ± {sd}"
    )
  ) |>
  modify_header(all_stat_cols() ~ "**{level}**") |>
  modify_footnote(everything() ~ NA) |>
  as_tibble() |>
  rename_with(~ str_remove_all(.x, "\\*")) |>
  filter(!if_any(everything(), is.na))
```




```{r demo_table}
tbl_summary(
  site_data,
  by = report_year,
  missing = "no",
  digits = list(all_continuous() ~ 1, all_categorical() ~ c(0, 1)),
  statistic = list(
    all_categorical() ~ "{n} ({p}%)",
    all_continuous() ~ "{mean} ± {sd}"
  )
) |>
  modify_header(all_stat_cols() ~ "**{level}**") |>
  modify_footnote(everything() ~ NA) |>
  as_tibble() |>
  rename_with(~ str_remove_all(.x, "\\*")) |>
  filter(!if_any(everything(), is.na)) |>
  # show popover for each cell in the table, with content from the reference data
  mutate(across(
    -Characteristic,
    ~ cell_spec(
      .x,
      "html",
      popover = spec_popover(
        title = "National Reference",
        content = ref_data[[cur_column()]],
        trigger = "hover"
      )
    )
  )) |>
  kbl(format = "html", booktabs = TRUE, escape = FALSE) |>
  pack_rows("Sex", 2, 4) |>
  pack_rows("ASA grade", 5, 8) |>
  pack_rows("Pre-fracture mobility", 9, 12) |>
  pack_rows("Fracture type", 13, 15) |>
  pack_rows("Residence", 16, 18) |>
  pack_rows("30-day mortality", 19, 20) |>
  pack_rows("365-day mortality", 21, 22)
```




# Rolling 3-year Risk-adjusted Mortality
  
::: {.callout-tip}
## Tip
Hover over the points in the funnel plot below to see the number of surgeries and risk-adjusted mortality rate for your site.
:::

## 30-day Mortality 

```{r 30d_mortality_data}

target_names <- tar_meta(store = here::here("_targets"))$name[
  str_detect(
    tar_meta(store = here::here("_targets"))$name,
    "mi_rollamr_mort30d"
  ) &
    str_detect(tar_meta(store = here::here("_targets"))$name, "au")
]
# Load them into the environment
tar_load(any_of(target_names), store = here::here("_targets"))

# Bundle them into a list
plot_list <- mget(target_names)

# Combine into a single data frame for plotting

dat <- plot_list |>
  bind_rows(.id = "frame") |>
  mutate(frame = str_remove(frame, "mi_rollamr_mort30d_")) |>
  mutate(frame = str_remove(frame, "_au")) |>
  mutate(report_id = factor(report_id))
```

```{r dynamic_funnel_function}
dynamic_funnel <- function(dat) {
  # Fix a per-site color and opacity up front so both persist across animation frames
  site_colors <- setNames(
    scales::hue_pal()(length(levels(dat$report_id))),
    levels(dat$report_id)
  )
  # Convert data to framed data for animation
  plot_dat <-
    expand.grid(frame = unique(dat$frame), report_id = levels(dat$report_id)) |>
    left_join(dat, by = c("frame", "report_id")) |>
    left_join(hoscode_data |> select(report_id, h_name), by = "report_id") |>
    mutate(no_data_flag = if_else(is.na(n), 1, 0)) |>
    arrange(frame) |>
    fill(smort, .by = report_id, .direction = "down") |>
    fill(
      ws_mean,
      ws_lcl95_exact,
      ws_ucl95_exact,
      ws_lcl99_exact,
      ws_ucl99_exact,
      .by = frame,
      .direction = "downup"
    ) |>
    mutate(across(where(is.numeric), ~ replace_na(.x, 0))) |>

    # Fix a per-site color and opacity up front so both persist across animation frames
    mutate(marker_color = site_colors[as.character(report_id)]) |>
    # Only display stats for the site of interest
    mutate(
      marker_text = if_else(
        report_id == params$site_id,
        paste0(
          h_name,
          " (",
          report_id,
          ")",
          "<br>Surgeries: ",
          n,
          "<br>Mortality: ",
          scales::percent(smort, accuracy = 0.1),
          " (Crude:",
          scales::percent(cmort, accuracy = 0.1),
          ")"
        ),
        ""
      ),
      marker_opacity = if_else(report_id == params$site_id, 1.0, 0.2),
    )

  # Foundation plotly skeleton

  fig <- plot_ly(
    plot_dat,
    x = ~n,
    y = ~smort,
    frame = ~frame,
    legendgroup = ~report_id,
    type = "scatter",
    mode = "markers",
    marker = list(
      size = 10,
      color = ~marker_color,
      opacity = ~marker_opacity
    ),
    text = ~marker_text,
    hoverinfo = "text",
    showlegend = FALSE
  ) |>
    layout(
      xaxis = list(title = "Rolling 3-year Surgical Volume", fixedrange = TRUE),
      yaxis = list(
        title = "Risk-adjusted Mortality Rate",
        fixedrange = TRUE,
        tickformat = ".0%"
      )
    ) |>
    animation_opts(
      frame = 2000,
      transition = 1000,
      easing = "cubic-in-out",
      redraw = TRUE
    ) |>
    animation_slider(
      currentvalue = list(
        prefix = "Year ",
        font = list(color = "rgba(4,66,133,1)")
      ),
      bgcolor = "rgba(6,126,195,1)",
      bordercolor = "rgba(4,66,133,1)",
      borderwidth = 2,
      activebgcolor = "rgba(4,66,133,1)"
    ) |>
    animation_button(
      bgcolor = "rgba(6,126,195,1)",
      borderwidth = 2,
      bordercolor = "rgba(4,66,133,1)"
    ) |>
    config(
      doubleClick = FALSE,
      displayModeBar = TRUE,
      displaylogo = FALSE,
      modeBarButtonsToRemove = c(
        "pan2d",
        "zoom2d",
        "select2d",
        "lasso2d",
        "autoScale2d",
        "resetScale2d",
        "zoomIn2d",
        "zoomOut2d"
      )
    )

  # Post rendering customisation

  fig <- plotly_build(fig)

  # Remove the "visible" attribute from each trace in each frame,
  # therefore selection of a report_id in one frame will be retained in the next frame
  fig$x$frames <- lapply(fig$x$frames, function(frame) {
    frame$data <- lapply(frame$data, function(trace) {
      trace$visible <- NULL
      trace
    })
    frame
  })

  # Redraw national mean and confidence interval for each frame,
  # taking into account of the report_id absent in that frame

  frame_ws_mean <- plot_dat |> distinct(frame, ws_mean) |> deframe()

  frame_curves <- plot_dat |>
    # n = 0 rows are filled-in placeholders (report_id absent that frame),
    # not real surgical volumes, so they'd distort the curve's start/end
    filter(no_data_flag == 0) |>
    select(
      frame,
      n,
      ws_lcl95_exact,
      ws_ucl95_exact,
      ws_lcl99_exact,
      ws_ucl99_exact
    ) |>
    arrange(frame, n) |>
    group_by(frame) |>
    group_split()
  names(frame_curves) <- sapply(frame_curves, function(d) d$frame[1])

  hline_shape <- function(y, color = "rgba(6,126,195,1)") {
    list(
      type = "line",
      x0 = 0,
      x1 = 1,
      xref = "paper",
      y0 = y,
      y1 = y,
      yref = "y",
      line = list(color = color, dash = "dash", width = 1.5)
    )
  }

  path_shape <- function(
    x,
    y,
    dash = "dot",
    width = 1,
    color = "rgba(6,126,195,1)"
  ) {
    path <- paste0(
      "M",
      x[1],
      ",",
      y[1],
      paste0(" L", x[-1], ",", y[-1], collapse = "")
    )
    list(
      type = "path",
      path = path,
      xref = "x",
      yref = "y",
      line = list(color = color, dash = dash, width = width)
    )
  }

  # Exact binomial control limit at an arbitrary n, using the frame's own ws_mean
  exact_ci_col <- function(ws_mean, n, alpha, col) {
    Hmisc::binconf(
      ws_mean * n,
      n,
      alpha = alpha,
      method = "exact",
      return.df = TRUE
    )[[col]]
  }

  # Each frame's curve only spans that frame's own range of n, so it stops
  # short of the shared, fixed x-axis edges. Extend it smoothly by evaluating
  # the exact binomial control-limit formula at extra n values between each
  # axis edge and the curve's first/last real point (rather than holding a
  # flat straight line at the endpoint), so the curve keeps following its
  # true shape all the way to both sides of the plot.
  extend_curve <- function(n, y, ws_mean, alpha, col, x_range, n_extra = 40) {
    left_n <- seq(max(x_range[1], 1), n[1], length.out = n_extra)
    right_n <- seq(n[length(n)], x_range[2], length.out = n_extra)
    left_y <- exact_ci_col(ws_mean, left_n, alpha, col)
    right_y <- exact_ci_col(ws_mean, right_n, alpha, col)
    list(
      x = c(left_n, n[-1], right_n[-1]),
      y = c(left_y, y[-1], right_y[-1])
    )
  }

  build_shapes_for_frame <- function(frame_name, x_range) {
    curves <- frame_curves[[frame_name]]
    ws_mean <- frame_ws_mean[[frame_name]]
    lcl95 <- extend_curve(
      curves$n,
      curves$ws_lcl95_exact,
      ws_mean,
      0.05,
      2,
      x_range
    )
    ucl95 <- extend_curve(
      curves$n,
      curves$ws_ucl95_exact,
      ws_mean,
      0.05,
      3,
      x_range
    )
    lcl99 <- extend_curve(
      curves$n,
      curves$ws_lcl99_exact,
      ws_mean,
      0.003,
      2,
      x_range
    )
    ucl99 <- extend_curve(
      curves$n,
      curves$ws_ucl99_exact,
      ws_mean,
      0.003,
      3,
      x_range
    )
    list(
      hline_shape(ws_mean),
      path_shape(lcl95$x, lcl95$y, dash = "dot", color = "rgba(0,115,0,1)"),
      path_shape(ucl95$x, ucl95$y, dash = "dot", color = "rgba(177,0,0,1)"),
      path_shape(
        lcl99$x,
        lcl99$y,
        dash = "dashdot",
        color = "rgba(0,115,0,1)"
      ),
      path_shape(
        ucl99$x,
        ucl99$y,
        dash = "dashdot",
        color = "rgba(177,0,0,1)"
      )
    )
  }

  mean_label_annotation <- function(frame_name) {
    y <- frame_ws_mean[[frame_name]]
    list(
      x = 1,
      xref = "paper",
      xanchor = "right",
      y = 1,
      yref = "paper",
      yanchor = "bottom",
      text = paste("National average:", scales::percent(y, accuracy = 0.1)),
      showarrow = FALSE,
      font = list(color = "rgba(4,66,133,1)", size = 20, bold = TRUE),
      yshift = -100
    )
  }

  first_frame_name <- fig$x$frames[[1]]$name
  axis_range <- fig$x$layout$xaxis$range
  fig$x$layout$shapes <- build_shapes_for_frame(first_frame_name, axis_range)
  fig$x$layout$annotations <- list(mean_label_annotation(first_frame_name))
  fig$x$frames <- lapply(fig$x$frames, function(fr) {
    fr$layout <- list(
      shapes = build_shapes_for_frame(fr$name, axis_range),
      annotations = list(mean_label_annotation(fr$name))
    )
    fr
  })

  return(fig)
}
```




```{r dynamic_funnel_30d}
#| fig-width: 8
#| fig-height: 8

dynamic_funnel(dat)
```



## 365-day Mortality

```{r 365d_mortality_data}

target_names <- tar_meta(store = here::here("_targets"))$name[
  str_detect(
    tar_meta(store = here::here("_targets"))$name,
    "mi_rollamr_mort365d"
  ) &
    str_detect(tar_meta(store = here::here("_targets"))$name, "au") &
    !str_detect(
      tar_meta(store = here::here("_targets"))$name,
      as.character(as.numeric(format(Sys.Date(), "%Y")) - 1)
    )
]
# Load them into the environment
tar_load(any_of(target_names), store = here::here("_targets"))

# Bundle them into a list
plot_list <- mget(target_names)

# Combine into a single data frame for plotting

dat <- plot_list |>
  bind_rows(.id = "frame") |>
  mutate(frame = str_remove(frame, "mi_rollamr_mort365d_")) |>
  mutate(frame = str_remove(frame, "_au")) |>
  mutate(report_id = factor(report_id))
```

```{r dynamic_funnel_365d}
#| fig-width: 8
#| fig-height: 8

dynamic_funnel(dat)
```

